Visual Diagnostics of a Model Explainer -- Tools for the Assessment of LIME Explanations

### 2020-07-15 paper.Rnw
added comment about where to add sentence about limeaid
start on limitations - overplotting

Throughout this paper, we make a distinction between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, {\it explanability} is defined as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

@@

The remainder of the paper is structured as follows. Section \ref{background} provides background and claims made by \citet{ribeiro:2016} about LIME. We introduce the suggested diagnostic plots in Section \ref{diagnostics}. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest model fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extension and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}.

@@

\section{Discussion} \label{discussion}

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own. We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots. Here we discuss the limitations and possible extensions of the diagnostic plots and concerns about LIME raise by applications of the diagnostics plots in this paper.

@@

% We acknowledge that not all prediction problems will produce trends as easy to interpret as the visualizations we have shown in the section, but we still believe that visualizations of the procedure can provide insight into the quality of the LIME explanation. We hope these figures will provide a starting point (if not an ending point) for the assessment of a single LIME explanation. Perhaps the inability to interpret a plot is itself telling, or the act of creating extensions of these visualizations to work with a different scenarios will lead to an understanding of the trustworthiness of the explanation.

>>>>

Throughout this paper, we distinguish between interpretability and explanability of models. We define {\it interpretability} as the ability to directly use model parameters to understand the relationships in the data captured by the model: e.g., a linear model coefficient associated with a predictor variable indicates the amount the response variable changes based on a change in the predictor variable. In contrast, {\it explanability} is defined as the ability to use the model in an indirect manner to understand the relationships in the data captured by the model: e.g., partial dependence plots depict the marginal relationship between model predictions and predictor variables \citep{friedman:2001}.

@@

The remainder of the paper is structured as follows. Section \ref{background} provides background and claims made by \citet{ribeiro:2016} about LIME. We introduce the suggested diagnostic plots in Section \ref{diagnostics}. Then in Section \ref{application}, we demonstrate the use of the diagnostics to assess LIME explanations for a random forest model fit to a forensics bullet matching dataset. Section \ref{discussion} concludes with a discussion on extension and limitations of the diagnostic plots and concerns about LIME in regards to the claims made by \citet{ribeiro:2016} brought about by the visualization examples in this paper that agree with \citet{alvarezmelis:2018}, \citet{laugel:2018}, and \citet{molnar:2019}. \kgc{Add sentence here about limeaid...}

@@

\section{Discussion} \label{discussion}
\hh{The discussion should not have subsection ... but we can keep them for now for structural purposes.}

This paper highlights that while an explainer model is meant to provide clarity, it actually adds another layer of complexity to predictive models by requiring yet another model that needs to be assessed. Without an assessment of the explainer model, LIME is a black-box procedure of its own. \hh{XXX what is the implication? bring it to a punch: asking the user to trust into explainer  ...} We suggest the use of visual diagnostics to counteract the black-box nature of LIME and provide three diagnostic plots. %Here we discuss the limitations and possible extensions of the diagnostic plots and concerns about LIME raise by applications of the diagnostics plots in this paper.

@@

The largest limitation to visualizations is the dimensionality of the data shown, both in the number of dimensions or features as well as the number of observations. 
Fortunately, in the situation of LIME, both of these aspects are rather well controlled: 
LIME relies heavily on simulations to generate data scenarios that are close to the data observed but exhibits variability.
Effects from overplotting should be relatively mild, because output from simulations is shown, which is expected to be (relatively) continuous. In that respect, the diagnostics shown for the \data\ and  in the bullet example  are representative of what is expected.
But in cases where overplotting does become problematic, the user could either simply reduce the size of the simulations or use some well-studied binning techniques in the visualizations, as discussed in \citet{Carr:1987hi} \hh{reference to Millions book}

% talking points: 
% We acknowledge that not all prediction problems will produce trends as easy to interpret as the visualizations we have shown in the section, but we still believe that visualizations of the procedure can provide insight into the quality of the LIME explanation. We hope these figures will provide a starting point (if not an ending point) for the assessment of a single LIME explanation. Perhaps the inability to interpret a plot is itself telling, or the act of creating extensions of these visualizations to work with a different scenarios will lead to an understanding of the trustworthiness of the explanation.
